---
title: "Penguins Assessment"
author: "Student"
date: "23/11/2022"
output: html_document
---

## Question 01: What is the reproducibility crisis in science and should we care?
300-500 words

## Question 02: Create a figure using the Palmer Penguin dataset that is correct but badly communicates the data.
Focus on visual elements rather than writing misleading text on it. Provide the figure and explain your choices. 100-200 words

```{r}

# Load packages
library(palmerpenguins)
library(ggplot2)
library(janitor)
library(dplyr)

# Set working directory
setwd("C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment") 

# Save data
write.csv(penguins_raw, "RawData/penguins_raw.csv") 
#inside 'RawData' folder in the working directory

# Clean data
source("C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment/Functions/Cleaning.r")
penguins_clean <- cleaning(penguins_raw)
# Cleaning function is saved in a separate r script, inside Functions folder

# Save clean data
write.csv(penguins_clean, "C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment/CleanData/clean_data.csv")

# View data
penguins_clean


```
## Creating a Figure

```{r}
# I am going to make a box plot of body masses for the two sexes

# First I extract sex and body mass, from the clean dataset and remove NAs for both
source("C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment/Functions/SubsetData.R")
sex_bodymass_data <- subset_bodymass(penguins_clean)
# The code for this is saved in the cleaning script in the functions folder

# Save 
write.csv(sex_bodymass_data, "C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment/CleanData/ sex_bodymass_data.csv") 

# Making the figure
bodymass_boxplot <- ggplot(data = sex_bodymass_data, 
                          aes(x = sex, 
                              y = body_mass_g)) +
                          geom_boxplot(aes(color = sex), show.legend = FALSE)
bodymass_boxplot

# Making it ugly
bad_boxplot <- ggplot(data = sex_bodymass_data, 
                          aes(x = sex, 
                              y = body_mass_g)) +
                          geom_boxplot(aes(color = sex, fill = sex), show.legend = FALSE) +
  geom_point(aes(size=10)) +
  scale_color_manual(values = c("green", "red")) +
  scale_fill_manual(values = c("red", "green"))
  

bad_boxplot
```
This box plot has two big problems.

1. Colours:

Though filling in the two boxes in separate colours is good practice, I have deliberately picked bad colours. The clash of red and green not only is hard on the eyes but makes viewing the mean difficult for people who are colourblind.

2. Data points:

The data points have been added using just geom_point() without using jitter. This means they all fall on the same position on the x-axis for each box and so it is impossible to pick out individual points. I have made the points extra large to emphasize this!


It would also be a good idea to re-label the x and y axes and to add a legend explaining where this data has come from and what it shows!



##Question 03: Discuss how colour choice is important for communication in science figures and what considerations should be made

##Question 04: Run a statistical test on the Palmer Penguins dataset and produce a figure to explain it.

I want to investigate whether island has an effect on the body mass of Adelie penguins.

There are three islands;	
Torgersen 	
Biscoe and 	
Dream

So I will need to perform a one-way-ANOVA 

```{r}
# Set working directory
setwd("C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment") 

# Load libraries - from Libraries.R in functions folder in working directory
source("C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment/Functions/Libraries.R")

# Load data
read.csv("RawData/penguins_raw.csv")

# Clean data
source("C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment/Functions/Cleaning.r")
penguins_clean <- cleaning(penguins_raw)
# Cleaning function is saved in a separate r script, inside Functions folder

# Save clean data
write.csv(penguins_clean, "C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment/CleanData/clean_data.csv")

# Extract data, remove NAs
source("C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment/Functions/Cleaning.r")
adelie_data <- subset_adelies(penguins_clean)
summary(adelie_data)

# Assumptions of ANOVA = independent data points, homogeneity of variances, normal distribution

# Equality of Variances - Levene test
library(car)
leveneTest(body_mass_g ~  island, data = adelie_data)
# non significant p value (p>0.05) means our variances are roughlt equal

# Normal Distribution - as the data set contains 151 samples (n>30), I will assume normality due to the central limit theorem

# ANOVA
penguins_anova <- aov(body_mass_g ~ island,
  data = adelie_data
)
summary(penguins_anova)
penguins_anova
hist(penguins_anova$residuals)
```

^
Here I took a subset of my data that only contained Adelie penguins, and the columns of body mass and island

I assumed normality because there was a big sample size. Then I did a Levene test which showed body mass had roughly have equal variances between islands. 

Then I was able to perform a one-way ANOVA to see the effect of island on body mass.

The resulting p-value was 0.968, so not a significant result. Body mass didn't differ significantly between islands.


# Making a figure
```{r}
# Set working directory
setwd("C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment") 

# Load libraries - from Libraries.R in functions folder in working directory
source("C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment/Functions/Libraries.R")

# Load data
read.csv("RawData/penguins_raw.csv")

# Clean data
source("C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment/Functions/Cleaning.r")
penguins_clean <- cleaning(penguins_raw)
# Cleaning function is saved in a separate r script, inside Functions folder

# Save clean data
write.csv(penguins_clean, "C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment/CleanData/clean_data.csv")

# Extract data, remove NAs
source("C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment/Functions/Cleaning.r")
adelie_data <- subset_adelies(penguins_clean)
summary(adelie_data)

# Figure
library(ggplot2)
adelies_figure <- ggplot(adelie_data) +
  aes(x = island, y = body_mass_g, color = island) +
   geom_boxplot(aes(colour=island), width = 0.4) +
  geom_jitter(aes(color = island), alpha = 0.3, show.legend = FALSE, position = position_jitter(width = 0.2, seed = 0)) +
  theme(legend.position = "none") +
  ggtitle("Body Mass in Adelies Does Not Vary Across Three Islands") +
  theme(plot.title = element_text(size=10)) +
  labs(x="Island", y= "Body Mass (g)") +
  theme_bw()
 


# Save figure
library(ragg)
agg_png("Figures/adelie_penguins.png", 
        width = 800, height = 600, units = "px",
        res = 300,
        scaling = 0.5)
adelies_figure
dev.off()


```

