---
title: "Q4"
author: "Student"
date: "05/12/2022"
output: html_document
---
##Question 04: Run a statistical test on the Palmer Penguins dataset and produce a figure to explain it.

I want to investigate whether island has an effect on the body mass of Adelie penguins.

There are three islands;	
Torgersen 	
Biscoe and 	
Dream

So I will need to perform a one-way-ANOVA 

```{r}
# Set working directory
setwd("C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment") 

# Load libraries - from Libraries.R in functions folder in working directory
source("C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment/Functions/Libraries.R")

# Load data
read.csv("RawData/penguins_raw.csv")

# Clean data
source("C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment/Functions/Cleaning.r")
penguins_clean <- cleaning(penguins_raw)
# Cleaning function is saved in a separate r script, inside Functions folder

# Save clean data
write.csv(penguins_clean, "C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment/CleanData/clean_data.csv")

# Extract data, remove NAs
source("C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment/Functions/Cleaning.r")
adelie_data <- subset_adelies(penguins_clean)
summary(adelie_data)

# Assumptions of ANOVA = independent data points, homogeneity of variances, normal distribution

# Equality of Variances - Levene test
library(car)
leveneTest(body_mass_g ~  island, data = adelie_data)
# non significant p value (p>0.05) means our variances are roughlt equal

# Normal Distribution - as the data set contains 151 samples (n>30), I will assume normality due to the central limit theorem

# ANOVA
penguins_anova <- aov(body_mass_g ~ island,
  data = adelie_data
)
summary(penguins_anova)
penguins_anova
hist(penguins_anova$residuals)
```
^
Here I took a subset of my data that only contained Adelie penguins, and the columns of body mass and island

I assumed normality because there was a big sample size. Then I did a Levene test which showed body mass had roughly have equal variances between islands. 

Then I was able to perform a one-way ANOVA to see the effect of island on body mass.

The resulting p-value was 0.968, so not a significant result. Body mass didn't differ significantly between islands.


# Making a figure
```{r}
# Set working directory
setwd("C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment") 

# Load libraries - from Libraries.R in functions folder in working directory
source("C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment/Functions/Libraries.R")

# Load data
read.csv("RawData/penguins_raw.csv")

# Clean data
source("C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment/Functions/Cleaning.r")
penguins_clean <- cleaning(penguins_raw)
# Cleaning function is saved in a separate r script, inside Functions folder

# Save clean data
write.csv(penguins_clean, "C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment/CleanData/clean_data.csv")

# Extract data, remove NAs
source("C:/Users/user/Desktop/Computing MTY3/PenguinsAssignment/Functions/Cleaning.r")
adelie_data <- subset_adelies(penguins_clean)
summary(adelie_data)

# Figure
library(ggplot2)
adelies_figure <- ggplot(adelie_data) +
  aes(x = island, y = body_mass_g, color = island) +
   geom_boxplot(aes(colour=island), width = 0.4) +
  geom_jitter(aes(color = island), alpha = 0.3, show.legend = FALSE, position = position_jitter(width = 0.2, seed = 0)) +
  theme(legend.position = "none") +
  ggtitle("Body Mass in Adelies Does Not Vary Across Three Islands") +
  theme(plot.title = element_text(size=10)) +
  labs(x="Island", y= "Body Mass (g)") +
  theme_bw()
 


# Save figure
library(ragg)
agg_png("adelie_penguins.png", 
        width = 800, height = 600, units = "px",
        res = 300,
        scaling = 0.5)
adelies_figure
dev.off()

adelies_figure


```